Incorporating periodic variability in hidden Markov models for animal movement

### 2016-02-29 paper3.Rnw
compiling problem, need to rewrite results for the paper
minor edits, cat14 has a small fitting problem

Adding temporal heterogeneity in the transitions for 3 to 6 state HSMM was a challenge because the number of parameters and model complexity increase dramatically, thus creating overfitting problems. Model reduction was used to balance out the increase in parameters and see if there are evidence suggesting adding temporal heterogeneity complexity can reduce the number of states to a statistically and biologically interpretable level. Without loss of generality, we assessed time homogeneous transition HMM models instead of HSMM and made sure each model reduction does not change the BIC selection of number of states, which HMM heavily depends on. Goodness of fit will decrease, but the information on the number of states remains the same.

@@

>>>>

Figure 1 shows the BIC-optimal number of states for time homogenous models is consistent with van de Kerk et al. Furthermore, BIC-optimal number of states does not depend on the model for both time-homogenous and time-inhomogenous within their respective framework, but time-inhomogenous models have the BIC-optimal number of states overall.

@@

<<BICred_plot cat131,warning=FALSE,echo=FALSE,message=FALSE,fig.cap="Change in BIC across models, cat \\#131.  $\\dbic$ is scaled individually per model, so that $\\dbic=0$ represents the BIC-optimal number of states for a particular model.",fig.width=6,fig.height=4>>=
load("cat2allmod.RData")
df2 <- transform(allmodsumdf,
            distrib=ifelse(grepl("LN",model),"log-Normal","Weibull"),
            turn_angle=ifelse(grepl("VM",model),"yes","no"),
            homog=ifelse(grepl("HMM",model),"homogeneous HMM","time-heterogeneous HMM"))
ggplot(df2, aes(x=nstates,y=deltaBIC,colour=distrib,lty=turn_angle))+
    geom_line()+
    facet_wrap(~ homog)+
    scale_colour_brewer(name="step length\ndistribution",palette="Set1")+
    ## scale_size_continuous( name="# parameters") +
    #    geom_point(aes(size=parameters))+
    scale_linetype(name="von Mises\nturning angle?")+
    labs(y=expression(Delta*"BIC"),
         x="Number of States")+
    geom_hline(yintercept=0,colour="gray")+
    zmargin
@
